---
title: "Anthropogenic CO2 from 1994 to 2007"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r parent, child = "/nfs/kryo/work/jenmueller/emlr_cant/utilities/setup_obs.Rmd"}
# this chunk runs the code stored in setup.Rmd
# if required, please refer to instructions given here:
# https://jdblischak.github.io/workflowr/articles/wflow-07-common-code.html
```

```{r define_paths, include=FALSE}
path_gruber_2019    <- "/nfs/kryo/work/updata/cant_gruber_2019/"
path_preprocessing  <- paste(path_root, "/observations/preprocessing/", sep = "")
```

```{r load_libraries_specific, include=FALSE}
library(tidync)
library(stars)
```


# Data source

- Anthropogenic CO2 estimates (1994-2007) by Gruber et al. (2019) downloaded in August 2020 from [NOAA/NCEI Ocean Carbon Data System (OCADS)](https://www.nodc.noaa.gov/ocads/oceans/ndp_100/ndp100.html){target="_blank"}

# Read nc files

Here, we use the standard case V101 for public and raw data sets.

## Public data sets

The publicly available data sets contain only positive Cant estimates.

### 3d fields

```{r read_Gruber_dcant}

# open file
dcant <- tidync(paste(
  path_gruber_2019,
  "dcant_emlr_cstar_gruber_94-07_vs1.nc",
  sep = ""
))

# read gamma field as tibble
dcant <- dcant %>%  activate(GAMMA_DENS)
dcant_gamma <- dcant %>% hyper_tibble()

# read delta cant field
dcant <- dcant %>%  activate(DCANT_01)
dcant <- dcant %>% hyper_tibble()

# join cant and gamma fields
dcant <- left_join(dcant, dcant_gamma)

# harmonize column names and coordinates
dcant <- dcant %>% 
  rename(lon = LONGITUDE,
         lat = LATITUDE,
         depth = DEPTH,
         gamma = GAMMA_DENS,
         dcant_pos = DCANT_01) %>% 
  mutate(lon = if_else(lon < 20, lon + 360, lon))

rm(dcant_gamma)

```

### Column inventories

```{r read_Gruber_dcant_inv_v101}

dcant_inv <- tidync(paste(
  path_gruber_2019,
  "inv_dcant_emlr_cstar_gruber_94-07_vs1.nc",
  sep = ""
))

dcant_inv <- dcant_inv %>%  activate(DCANT_INV01)
dcant_inv <- dcant_inv %>% hyper_tibble()

# harmonize column names and coordinates
dcant_inv <- dcant_inv %>% 
  rename(lon = LONGITUDE,
         lat = LATITUDE,
         dcant_pos = DCANT_INV01) %>% 
  mutate(lon = if_else(lon < 20, lon + 360, lon))

```

```{r read_Gruber_dcant_inv_all}

dcant_inv_all <- read_ncdf(
  paste(
    path_gruber_2019,
    "inv_dcant_emlr_cstar_gruber_94-07_vs1.nc",
    sep = ""
  ),
  var = sprintf("DCANT_INV%02d", seq(1, 14, 1)),
  make_units = FALSE
)

dcant_inv_all <- dcant_inv_all %>% as_tibble()

dcant_inv_all <- dcant_inv_all %>% 
  pivot_longer(DCANT_INV01:DCANT_INV14,
               names_to = "Version_ID",
               values_to = "dcant_pos",
               names_prefix = "DCANT_INV")

# harmonize column names and coordinates
dcant_inv_all <- dcant_inv_all %>%
  rename(lon = LONGITUDE,
         lat = LATITUDE) %>%
  mutate(lon = if_else(lon < 20, lon + 360, lon))

```

## Raw data

Internally available data sets also contain negative Cant estimates, as they are generated in the "raw" output of the eMLR mapping step.

```{r read_Gruber_V101}

# open v 101 file
V101 <- tidync(paste(path_gruber_2019,
                     "Cant_V101new.nc",
                     sep = ""))

# create tibble
V101 <- V101 %>%  activate(Cant)
V101 <- V101 %>% hyper_tibble()

# harmonize column names and coordinates
V101 <- V101 %>% 
  rename(lon = longitude,
         lat = latitude,
         dcant = Cant) %>% 
  filter(dcant != -999) %>% 
  mutate(lon = if_else(lon < 20, lon + 360, lon))

```


# Apply basin mask

```{r apply_basin_mask}

# use only three basin to assign general basin mask
# ie this is not specific to the MLR fitting
basinmask <- basinmask %>% 
  filter(MLR_basins == "2") %>% 
  select(lat, lon, basin_AIP)

dcant <- inner_join(dcant, basinmask)
dcant_inv_publ <- inner_join(dcant_inv, basinmask)
dcant_inv_all <- inner_join(dcant_inv_all, basinmask)
V101 <- inner_join(V101, basinmask)

```

# Join pos and all delta Cant

```{r join_V101_and_all}

# join files
dcant_3d <- inner_join(dcant, V101)

rm(dcant, V101)

```

# Zonal mean section

```{r calculate_zonal_mean_section}

dcant_zonal <- m_zonal_mean_sd(dcant_3d)

```

# Column inventory 

## Calculation

```{r calculate_column_inventory}

dcant_inv_layers <- m_dcant_inv(dcant_3d)

dcant_inv <- dcant_inv_layers %>% 
  filter(inv_depth == params_global$inventory_depth_standard)

```

## Plots

### All Cant

```{r G19_cant_inv_maps, fig.asp=0.6}

p_map_cant_inv(
  df = dcant_inv,
  var = "dcant",
  col = "divergent")

```

### Pos Cant

```{r G19_cant_pos_inventory_maps, fig.asp=0.6}

p_map_cant_inv(
  df = dcant_inv,
  var = "dcant_pos")

```

```{r G19_cant_pos_inventory_maps_all, fig.asp=1.5}

p_map_cant_inv(
  df = dcant_inv_all %>% mutate(dcant_pos = dcant_pos*(10/13)),
  var = "dcant_pos") +
  facet_wrap(~ Version_ID, ncol = 2)

```

### Published inventories

```{r G19_cant_pos_inventory_maps_published, fig.asp=0.6}

p_map_cant_inv(
  df = dcant_inv,
  var = "dcant_pos")

```

### Published vs calculated

```{r G19_cant_offset_inventory, fig.asp=0.6}

# join published and calculated data sets
dcant_inv_offset <- inner_join(
  dcant_inv %>% rename(dcant_re = dcant_pos),
  dcant_inv_publ %>% rename(dcant_pub = dcant_pos)
)

# calculate offset
dcant_inv_offset <- dcant_inv_offset %>% 
  mutate(dcant_offset = dcant_re - dcant_pub)

# plot map
p_map_cant_inv(
  df = dcant_inv_offset,
  var = "dcant_offset",
  col = "divergent",
  breaks = seq(-3, 3, 0.25)
)

rm(dcant_inv_offset, dcant_inv_publ)

```


# Horizontal plane maps

## All Cant

```{r Cant_maps, fig.asp=0.6}

p_map_climatology(
  df = dcant_3d,
  var = "dcant",
  col = "divergent")

```

## Positive Cant

```{r Cant_maps_V101, fig.asp=0.6}

p_map_climatology(
  df = dcant_3d,
  var = "dcant_pos")

```

## Neutral density

```{r Gamma_maps_V101, fig.asp=0.6}

p_map_climatology(
  df = dcant_3d,
  var = "gamma")

```


# Zonal mean section plot

## Positive Cant

```{r Cant_pos_zonal_mean_sections}

dcant_zonal %>%
  group_split(basin_AIP) %>%
  head(1) %>%
  map(
    ~ p_section_zonal(
      df = .x,
      var = "dcant_pos_mean",
      plot_slabs = "n",
      subtitle_text = paste("Basin:", unique(.x$basin_AIP))
    )
  )

```

# Global sections plot

## All Cant

```{r Cant_sections_V101}

p_section_global(
  df = dcant_3d,
  var = "dcant",
  col = "divergent")

```

## Positive Cant

```{r Cant_sections}

p_section_global(
  df = dcant_3d,
  var = "dcant_pos")

```


# Sections at regular longitudes

## All Cant

```{r Cant_sections_regular, fig.asp=1.7}

p_section_climatology_regular(
  df = dcant_3d,
  var = "dcant",
  col = "divergent")

```

## Positive Cant

```{r Cant_sections_regular_V101, fig.asp=1.7}

p_section_climatology_regular(
  df = dcant_3d,
  var = "dcant_pos")

```

## Neutral density

```{r Gamma_sections_regular_V101, fig.asp=1.7}

p_section_climatology_regular(
  df = dcant_3d,
  var = "gamma")

```

# Write files

```{r write_Gruber_Cant_file}

dcant_3d %>%
  write_csv(paste(path_preprocessing,
                  "G19_dcant_3d.csv",
                  sep = ""))

dcant_inv %>%
  write_csv(paste(path_preprocessing,
                  "G19_dcant_inv.csv",
                  sep = ""))

dcant_inv_all %>%
  write_csv(paste(path_preprocessing,
                  "G19_dcant_inv_all.csv",
                  sep = ""))

dcant_zonal %>%
  write_csv(paste(path_preprocessing,
                  "G19_dcant_zonal.csv",
                  sep = ""))

```


# RECCAP2-ocean


```{r format_Gruber_dcant_for_RECCAP2}

# extract coordinate reference system
G19_raster <- raster::brick(paste0(
  path_gruber_2019,
  "dcant_emlr_cstar_gruber_94-07_vs1.nc"))

coord_ref <- raster::crs(G19_raster)
rm(G19_raster)

# open nc file for data extraction
dcant_nc <- tidync(paste(
  path_gruber_2019,
  "dcant_emlr_cstar_gruber_94-07_vs1.nc",
  sep = ""
))

# read delta cant field
dcant <- dcant_nc %>%
  activate(DCANT_01) %>%
  hyper_tibble(na.rm = FALSE)

# read delta cant field
gamma <- dcant_nc %>%
  activate(GAMMA_DENS) %>%
  hyper_tibble(na.rm = FALSE)

# join gamma and dcant
dcant <- full_join(dcant, gamma)
rm(gamma)

# harmonize column names and coordinates
dcant <- dcant %>% 
  rename(lon = LONGITUDE,
         lat = LATITUDE,
         depth = DEPTH,
         dcant = DCANT_01,
         gamma = GAMMA_DENS) %>% 
  mutate(gamma = if_else(is.na(dcant), NaN, gamma))

# convert dcant unit from "µmol kg-1" to "mol m-3"
dcant <- dcant %>% 
  mutate(dens = (1000 + gamma) / 1000,
         dcant = dcant * dens * 1e-3)

# create volume grid
dcant <- dcant %>% 
  m_layer_thickness() %>% 
  mutate(surface_area = marelac::earth_surf(lat, lon),
         volume = layer_thickness * surface_area,
         volume = if_else(is.na(dcant), NaN, volume))

# check total volume
dcant %>% 
  summarise(total_ocean_volume = sum(volume, na.rm = TRUE))

# check total dcant
dcant %>% 
  filter(depth <= 3000) %>% 
  mutate(dcant_inv = dcant * volume) %>% 
  summarise(total_dcant = sum(dcant_inv, na.rm = TRUE)*12*1e-15)

# select relevant columns
dcant <- dcant %>% 
  select(lon, lat, depth, dcant, volume)

# create raster objects
volume_raster <- dcant %>% 
  select(lon, lat, volume) %>%
  base::split(dcant$depth) %>% 
  lapply(raster::rasterFromXYZ) %>% 
  raster::brick() %>% 
  raster::setZ(z = unique(dcant$depth), name = "depth")

dcant_raster <- dcant %>% 
  select(lon, lat, dcant) %>%
  base::split(dcant$depth) %>% 
  lapply(raster::rasterFromXYZ) %>% 
  raster::brick() %>% 
  raster::setZ(z = unique(dcant$depth), name = "depth")

# assign coordinate reference system
raster::crs(dcant_raster) <- coord_ref
raster::crs(volume_raster) <- coord_ref

# assign NA values
raster::NAvalue(dcant_raster) <- -9999
raster::NAvalue(dcant_raster)
raster::NAvalue(volume_raster) <- -9999
raster::NAvalue(volume_raster)

# check object
dim(dcant_raster)
raster::nbands(dcant_raster)
raster::nlayers(dcant_raster)
names(dcant_raster) #get the names of layers
raster::getZ(dcant_raster)


# write netcdf file
raster::writeRaster(
  dcant_raster,
  filename = paste0(path_preprocessing,
                    "dcant_Gruber2019_1994-2007_v20211012.nc"),
  overwrite = T
)

raster::writeRaster(
  volume_raster,
  filename = paste0(path_preprocessing,
                    "volume_Gruber2019_1994-2007_v20211012.nc"),
  overwrite = T
)


# modify created netcdf files
library(ncdf4)

# dcant file

# open file in writing mode
dcant_reopen <- nc_open(
  paste0(path_preprocessing,
         "dcant_Gruber2019_1994-2007_v20211012.nc"),
  write = TRUE)

dcant_reopen
print(dcant_reopen)
names(dcant_reopen$var)

# change variable name
ncvar_rename(dcant_reopen, "depth", "dcant")

# close and reopen
nc_close(dcant_reopen)
dcant_reopen <- nc_open(
  paste0(path_preprocessing,
         "dcant_Gruber2019_1994-2007_v20211012.nc"),
  write = TRUE)

# add units
ncatt_get(dcant_reopen, varid = "dcant")
ncatt_put(dcant_reopen, varid = "dcant",
          attname = "units", attval = "mol m-3")

ncatt_get(dcant_reopen, varid = "z")
ncatt_put(dcant_reopen, varid = "z",
          attname = "units", attval = "metres")

nc_close(dcant_reopen)


# volume file

# open file in writing mode
volume_reopen <- nc_open(
  paste0(path_preprocessing,
         "volume_Gruber2019_1994-2007_v20211012.nc"),
  write = TRUE)

volume_reopen
print(volume_reopen)
names(volume_reopen$var)

# change variable name
ncvar_rename(volume_reopen, "depth", "volume")

# close and reopen
nc_close(volume_reopen)
volume_reopen <- nc_open(
  paste0(path_preprocessing,
         "volume_Gruber2019_1994-2007_v20211012.nc"),
  write = TRUE)

# add units
ncatt_get(volume_reopen, varid = "volume")
ncatt_put(volume_reopen, varid = "volume",
          attname = "units", attval = "m3")

ncatt_get(volume_reopen, varid = "z")
ncatt_put(volume_reopen, varid = "z",
          attname = "units", attval = "metres")

nc_close(volume_reopen)



# final check dcant

dcant_reopen <- tidync(
  paste0(path_preprocessing,
         "dcant_Gruber2019_1994-2007_v20211012.nc")) %>% 
  hyper_tibble()

dcant_reopen %>% 
  filter(z == 0) %>% 
  ggplot(aes(longitude, latitude, fill=dcant)) +
  geom_raster() +
  scale_fill_viridis_c()

dcant_reopen %>% 
  filter(longitude == 200.5) %>% 
  ggplot(aes(latitude, z, z=dcant)) +
  scale_y_reverse() +
  geom_contour_filled() +
  scale_fill_viridis_d()


dcant_reopen <- read_ncdf(
  paste0(path_preprocessing,
         "dcant_Gruber2019_1994-2007_v20211012.nc"))

plot(dcant_reopen,
     axes = TRUE)


# final check volume

volume_reopen <- tidync(
  paste0(path_preprocessing,
         "volume_Gruber2019_1994-2007_v20211012.nc")) %>% 
  hyper_tibble()

volume_reopen %>% 
  filter(z == 0) %>% 
  ggplot(aes(longitude, latitude, fill=volume)) +
  geom_raster() +
  scale_fill_viridis_c()

volume_reopen %>% 
  filter(longitude == 200.5) %>% 
  ggplot(aes(latitude, z, z=volume)) +
  scale_y_reverse() +
  geom_contour_filled() +
  scale_fill_viridis_d()


volume_reopen <- read_ncdf(
  paste0(path_preprocessing,
         "volume_Gruber2019_1994-2007_v20211012.nc"))

plot(volume_reopen,
     axes = TRUE)

```



