---
title: "Anthropogenic CO2 from 1994 to 2007"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r parent, child = "/nfs/kryo/work/jenmueller/emlr_cant/utilities/setup.Rmd"}
# this chunk runs the code stored in setup.Rmd
# if required, please refer to instructions given here:
# https://jdblischak.github.io/workflowr/articles/wflow-07-common-code.html
```

```{r define_paths, include=FALSE}
path_gruber_2019    <- "/nfs/kryo/work/updata/cant_gruber_2019/"
path_preprocessing  <- paste(path_root, "/observations/preprocessing/", sep = "")
```

```{r load_libraries_specific, include=FALSE}
library(tidync)
library(stars)
```


# Data source

- Anthropogenic CO2 estimates (1994-2007) by Gruber et al. (2019) downloaded in August 2020 from [NOAA/NCEI Ocean Carbon Data System (OCADS)](https://www.nodc.noaa.gov/ocads/oceans/ndp_100/ndp100.html){target="_blank"}

# Read nc files

Here, we use the standard case V101 for public and raw data sets.

## Public data sets

The publicly available data sets contain only positive Cant estimates.

### 3d fields

```{r read_Gruber_dcant}

# open file
dcant <- tidync(paste(
  path_gruber_2019,
  "dcant_emlr_cstar_gruber_94-07_vs1.nc",
  sep = ""
))

# read gamma field as tibble
dcant <- dcant %>%  activate(GAMMA_DENS)
dcant_gamma <- dcant %>% hyper_tibble()

# read delta cant field
dcant <- dcant %>%  activate(DCANT_01)
dcant <- dcant %>% hyper_tibble()

# join cant and gamma fields
dcant <- left_join(dcant, dcant_gamma)

# harmonize column names and coordinates
dcant <- dcant %>% 
  rename(lon = LONGITUDE,
         lat = LATITUDE,
         depth = DEPTH,
         gamma = GAMMA_DENS,
         dcant_pos = DCANT_01)

dcant_3d_reccap <- dcant

dcant <- dcant %>% 
  mutate(lon = if_else(lon < 20, lon + 360, lon))

rm(dcant_gamma)

```

### Column inventories

```{r read_Gruber_dcant_inv_v101}

dcant_inv_publ <- tidync(paste(
  path_gruber_2019,
  "inv_dcant_emlr_cstar_gruber_94-07_vs1.nc",
  sep = ""
))

dcant_inv_publ <- dcant_inv_publ %>%  activate(DCANT_INV01)
dcant_inv_publ <- dcant_inv_publ %>% hyper_tibble()

# harmonize column names and coordinates
dcant_inv_publ <- dcant_inv_publ %>% 
  rename(lon = LONGITUDE,
         lat = LATITUDE,
         dcant_pos = DCANT_INV01) %>% 
  mutate(lon = if_else(lon < 20, lon + 360, lon))

```

```{r read_Gruber_dcant_inv_all}

dcant_inv_publ_all <- read_ncdf(
  paste(
    path_gruber_2019,
    "inv_dcant_emlr_cstar_gruber_94-07_vs1.nc",
    sep = ""
  ),
  var = sprintf("DCANT_INV%02d", seq(1, 14, 1)),
  make_units = FALSE
)

dcant_inv_publ_all <- dcant_inv_publ_all %>% as_tibble()

dcant_inv_publ_all <- dcant_inv_publ_all %>% 
  pivot_longer(DCANT_INV01:DCANT_INV14,
               names_to = "Version_ID",
               values_to = "dcant_pos",
               names_prefix = "DCANT_INV")

# harmonize column names and coordinates
dcant_inv_publ_all <- dcant_inv_publ_all %>%
  rename(lon = LONGITUDE,
         lat = LATITUDE) %>%
  mutate(lon = if_else(lon < 20, lon + 360, lon))

```

## Raw data

Internally available data sets also contain negative Cant estimates, as they are generated in the "raw" output of the eMLR mapping step.

```{r read_Gruber_V101}

# open v 101 file
V101 <- tidync(paste(path_gruber_2019,
                     "Cant_V101new.nc",
                     sep = ""))

# create tibble
V101 <- V101 %>%  activate(Cant)
V101 <- V101 %>% hyper_tibble()

# harmonize column names and coordinates
V101 <- V101 %>% 
  rename(lon = longitude,
         lat = latitude,
         dcant = Cant) %>% 
  filter(dcant != -999) %>% 
  mutate(lon = if_else(lon < 20, lon + 360, lon))

```


# Apply basin mask

```{r apply_basin_mask}

# use only three basin to assign general basin mask
# ie this is not specific to the MLR fitting
basinmask <- basinmask %>% 
  filter(MLR_basins == "2") %>% 
  select(lat, lon, basin_AIP)

dcant <- inner_join(dcant, basinmask)
dcant_inv_publ_masked <- inner_join(dcant_inv_publ, basinmask)
dcant_inv_publ_all <- inner_join(dcant_inv_publ_all, basinmask)
V101 <- inner_join(V101, basinmask)

ggplot() +
  geom_tile(data = dcant_inv_publ,
            aes(lon, lat, fill = "basin mask not applied")) +
  geom_tile(data = dcant_inv_publ_masked,
            aes(lon, lat, fill = "basin mask applied")) +
  coord_quickmap()


```

# Join pos and all delta Cant

```{r join_V101_and_all}

# join files
dcant_3d <- inner_join(dcant, V101)

rm(dcant, V101)

```

# Zonal mean section

```{r calculate_zonal_mean_section}

dcant_zonal <- m_zonal_mean_sd(dcant_3d)

```

# Column inventory 

## Calculation

```{r calculate_column_inventory}

dcant_inv_layers <- m_dcant_inv(dcant_3d)

dcant_inv <- dcant_inv_layers %>% 
  filter(inv_depth == params_global$inventory_depth_standard)

```

## Plots

### All Cant

```{r G19_cant_inv_maps, fig.asp=0.6}

p_map_cant_inv(
  df = dcant_inv,
  var = "dcant",
  col = "divergent")

```

### Pos Cant

```{r G19_cant_pos_inventory_maps, fig.asp=0.6}

p_map_cant_inv(
  df = dcant_inv,
  var = "dcant_pos")

```

```{r G19_cant_pos_inventory_maps_all, fig.asp=1.5}

p_map_cant_inv(
  df = dcant_inv_publ_all %>% mutate(dcant_pos = dcant_pos*(10/13)),
  var = "dcant_pos") +
  facet_wrap(~ Version_ID, ncol = 2)

```

### Published inventories

```{r G19_cant_pos_inventory_maps_published, fig.asp=0.6}

p_map_cant_inv(
  df = dcant_inv_publ,
  var = "dcant_pos",
  title_text = "Published column inventories - unmasked")

p_map_cant_inv(
  df = dcant_inv_publ_masked,
  var = "dcant_pos",
  title_text = "Published column inventories - masked")

```

### Published vs calculated

```{r G19_cant_offset_inventory, fig.asp=0.6}

# join published and calculated data sets
dcant_inv_offset <- inner_join(
  dcant_inv %>% rename(dcant_re = dcant_pos),
  dcant_inv_publ_masked %>% rename(dcant_pub = dcant_pos)
)

# calculate offset
dcant_inv_offset <- dcant_inv_offset %>% 
  mutate(dcant_offset = dcant_re - dcant_pub)

# plot map
p_map_cant_inv(
  df = dcant_inv_offset,
  var = "dcant_offset",
  col = "bias",
  breaks = seq(-3, 3, 0.25)
)

rm(dcant_inv_offset)

```


# Horizontal plane maps

## All Cant

```{r Cant_maps, fig.asp=0.6}

p_map_climatology(
  df = dcant_3d,
  var = "dcant",
  col = "divergent")

```

## Positive Cant

```{r Cant_maps_V101, fig.asp=0.6}

p_map_climatology(
  df = dcant_3d,
  var = "dcant_pos")

```

## Neutral density

```{r Gamma_maps_V101, fig.asp=0.6}

p_map_climatology(
  df = dcant_3d,
  var = "gamma")

```


# Zonal mean section plot

## Positive Cant

```{r Cant_pos_zonal_mean_sections}

dcant_zonal %>%
  group_split(basin_AIP) %>%
  # head(1) %>%
  map(
    ~ p_section_zonal(
      df = .x,
      var = "dcant_pos_mean",
      plot_slabs = "n",
      subtitle_text = paste("Basin:", unique(.x$basin_AIP))
    )
  )

```

# Global sections plot

## All Cant

```{r Cant_sections_V101}

p_section_global(
  df = dcant_3d,
  var = "dcant",
  col = "divergent")

```

## Positive Cant

```{r Cant_sections}

p_section_global(
  df = dcant_3d,
  var = "dcant_pos")

```


# Sections at regular longitudes

## All Cant

```{r Cant_sections_regular, fig.asp=1.7}

p_section_climatology_regular(
  df = dcant_3d,
  var = "dcant",
  col = "divergent")

```

## Positive Cant

```{r Cant_sections_regular_V101, fig.asp=1.7}

p_section_climatology_regular(
  df = dcant_3d,
  var = "dcant_pos")

```

## Neutral density

```{r Gamma_sections_regular_V101, fig.asp=1.7}

p_section_climatology_regular(
  df = dcant_3d,
  var = "gamma")

```

# Write files

```{r write_Gruber_Cant_file}

dcant_3d %>%
  write_csv(paste(path_preprocessing,
                  "G19_dcant_3d.csv",
                  sep = ""))

dcant_inv %>%
  write_csv(paste(path_preprocessing,
                  "G19_dcant_inv.csv",
                  sep = ""))

dcant_inv_publ %>%
  write_csv(paste(path_preprocessing,
                  "G19_dcant_inv_publ.csv",
                  sep = ""))

dcant_inv_publ_all %>%
  write_csv(paste(path_preprocessing,
                  "G19_dcant_inv_all.csv",
                  sep = ""))

dcant_zonal %>%
  write_csv(paste(path_preprocessing,
                  "G19_dcant_zonal.csv",
                  sep = ""))

```


# RECCAP2-ocean submission

## Read WOA18

```{r read_WOA18}


path_woa2018 <- "/nfs/kryo/work/updata/woa2018/"

WOA18 <- tidync(paste(
  path_woa2018,
  "temperature/decav/1.00/woa18_decav_t00_01.nc",
  sep = ""
))

WOA18 <- WOA18 %>%
  hyper_tibble()

WOA18 <- WOA18 %>%
  distinct(lat, lon, depth)

WOA18 <- WOA18 %>%
  mutate(lon = if_else(lon < 0, lon + 360, lon))

```


## Read RECCAP2 biome mask

```{r read_RECCAP2_basin_mask}


path_basin_mask <-
  "/nfs/kryo/work/updata/reccap2/"

region_masks_all <-
  read_ncdf(paste(
    path_basin_mask,
    "RECCAP2_region_masks_all_v20220620.nc",
    sep = ""
  )) %>%
  as_tibble()

region_masks_all <- region_masks_all %>%
  select(-seamask)

region_masks_all <- region_masks_all %>%
  pivot_longer(open_ocean:southern,
               names_to = "region",
               values_to = "value") %>%
  mutate(value = as.factor(value))

region_masks_all %>%
  filter(value != 0,
         region == "atlantic") %>%
  ggplot(aes(lon, lat, fill = region)) +
  geom_raster() +
  scale_fill_brewer(palette = "Dark2") +
  coord_quickmap(expand = 0)

region_masks_atlantic <-
  region_masks_all %>%
  filter(value != 0,
         region == "atlantic")

region_masks_open_ocean <-
  region_masks_all %>%
  filter(value != 0,
         !(region %in% c("open_ocean", "arctic")))

rm(region_masks_all)

```

## Extensions to N-Atlantic

```{r extension_N_Atlantic}

dcant_3d <- dcant_3d_reccap %>% 
  filter(depth <= 3000)

ggplot() +
  geom_tile(data = region_masks_atlantic,
            aes(lon, lat, fill = region), alpha = 0.6) +
  geom_tile(data = dcant_3d %>% distinct(lon, lat),
            aes(lon, lat, fill = "G19"), alpha = 0.6) +
  scale_fill_brewer(palette = "Dark2") +
  coord_quickmap(expand = 0)

ggplot() +
  geom_tile(data = region_masks_open_ocean,
            aes(lon, lat, fill = "region"), alpha = 0.6) +
  geom_tile(data = dcant_3d %>% distinct(lon, lat),
            aes(lon, lat, fill = "G19"), alpha = 0.6) +
  scale_fill_brewer(palette = "Dark2") +
  coord_quickmap(expand = 0)

dcant_3d <- inner_join(dcant_3d,
                       region_masks_open_ocean %>% distinct(lon, lat))


ggplot() +
  geom_tile(data = region_masks_atlantic %>% filter(lat > 64),
            aes(lon, lat, fill = region), alpha = 0.6) +
  geom_tile(data = dcant_3d %>% distinct(lon, lat),
            aes(lon, lat, fill = "G19"), alpha = 0.6) +
  scale_fill_brewer(palette = "Dark2") +
  coord_quickmap(expand = 0)

ggplot() +
  geom_tile(data = region_masks_atlantic %>% filter(lat > 64),
            aes(lon, lat, fill = region), alpha = 0.6) +
  geom_tile(data = WOA18 %>% distinct(lon, lat),
            aes(lon, lat, fill = "WOA18"), alpha = 0.6) +
  scale_fill_brewer(palette = "Dark2") +
  coord_quickmap(expand = 0)

NA_basin_WOA18 <- left_join(
  region_masks_atlantic %>% filter(lat > 62) %>% distinct(lon, lat),
  WOA18
)


ggplot() +
  geom_tile(data = region_masks_atlantic,
            aes(lon, lat),
            alpha = 0.6) +
  geom_tile(
    data = NA_basin_WOA18 %>%
      group_by(lon, lat) %>%
      summarise(depth = max(depth)) %>%
      ungroup(),
    aes(lon, lat, fill = depth)
  ) +
  scale_fill_viridis_c(direction = -1) +
  coord_quickmap(expand = 0)

NA_basin_WOA18 <- NA_basin_WOA18 %>% 
  filter(depth %in% unique(dcant_3d$depth))

dcant_3d_boundary <- inner_join(
  NA_basin_WOA18,
  dcant_3d
)

dcant_3d_boundary %>% 
  ggplot(aes(dcant_pos, depth, col = lon)) +
  geom_point() +
  scale_color_viridis_c() +
  scale_y_reverse()

dcant_3d_boundary %>% 
  ggplot(aes(gamma, depth, col = lon)) +
  geom_point() +
  scale_color_viridis_c() +
  scale_y_reverse()

dcant_boundary_profile <- dcant_3d_boundary %>% 
  group_by(depth) %>% 
  summarise(dcant_pos = mean(dcant_pos, na.rm = TRUE),
            gamma = mean(gamma, na.rm = TRUE)) %>% 
  ungroup()

dcant_boundary_profile %>% 
  ggplot(aes(dcant_pos, depth)) +
  geom_point() +
  scale_color_viridis_c() +
  scale_y_reverse()

dcant_boundary_profile %>% 
  ggplot(aes(gamma, depth)) +
  geom_point() +
  scale_color_viridis_c() +
  scale_y_reverse()

dcant_boundary_profile <- dcant_boundary_profile %>% 
  mutate(dcant_pos = if_else(depth >= 2000, min(dcant_pos), dcant_pos))

NA_basin_WOA18 <- full_join(NA_basin_WOA18 %>% filter(lat > 64),
                            dcant_boundary_profile)

dcant_3d <-
  bind_rows(dcant_3d,
            NA_basin_WOA18 %>% filter(lat > 65))

```


# Column inventory 

## Calculation

```{r calculate_column_inventory_G19_filled}

dcant_inv_layers <- m_dcant_inv(dcant_3d %>% mutate(dcant = 0,
                                                    basin_AIP = "global"))

dcant_inv <- dcant_inv_layers %>% 
  filter(inv_depth == params_global$inventory_depth_standard)

```

## Map


```{r G19_cant_inv_maps_filled, fig.asp=0.6}

dcant_inv %>% 
  ggplot(aes(lon, lat, fill = dcant_pos)) +
  geom_tile() +
  scale_fill_viridis_c() +
  coord_quickmap(expand = 0)

```




```{r format_Gruber_dcant_for_RECCAP2}

# extract coordinate reference system
G19_raster <- raster::brick(paste0(
  path_gruber_2019,
  "dcant_emlr_cstar_gruber_94-07_vs1.nc"))

coord_ref <- raster::crs(G19_raster)
rm(G19_raster)

# open nc file for data extraction
dcant_nc <- tidync(paste(
  path_gruber_2019,
  "dcant_emlr_cstar_gruber_94-07_vs1.nc",
  sep = ""
))

# read delta cant field
dcant <- dcant_nc %>%
  activate(DCANT_01) %>%
  hyper_tibble(na.rm = FALSE)

# read delta cant field
gamma <- dcant_nc %>%
  activate(GAMMA_DENS) %>%
  hyper_tibble(na.rm = FALSE)

# join gamma and dcant
dcant <- full_join(dcant, gamma)
rm(gamma)

# harmonize column names and coordinates
dcant <- dcant %>%
  rename(lon = LONGITUDE,
         lat = LATITUDE,
         depth = DEPTH,
         dcant = DCANT_01,
         gamma = GAMMA_DENS) %>%
  mutate(gamma = if_else(is.na(dcant), NaN, gamma))

dcant <-
  full_join(dcant_3d,
            dcant %>% distinct(lat, lon, depth))


# convert dcant unit from "µmol kg-1" to "mol m-3"
dcant <- dcant %>% 
  rename(dcant = dcant_pos) %>%
  mutate(dens = (1000 + gamma) / 1000,
         dcant = dcant * dens * 1e-3)

# create volume grid
dcant <- dcant %>% 
  m_layer_thickness() %>% 
  mutate(surface_area = marelac::earth_surf(lat, lon),
         volume = layer_thickness * surface_area,
         volume = if_else(is.na(dcant), NaN, volume))

# check total volume
dcant %>%
  filter(depth <= 3000) %>%
  summarise(total_ocean_volume = sum(volume, na.rm = TRUE))

# check total dcant
dcant %>% 
  filter(depth <= 3000) %>%
  mutate(dcant_inv = dcant * volume) %>% 
  summarise(total_dcant = sum(dcant_inv, na.rm = TRUE)*12*1e-15)

# select relevant columns
dcant <- dcant %>% 
  select(lon, lat, depth, dcant, volume)

# create raster objects
volume_raster <- dcant %>% 
  select(lon, lat, volume) %>%
  base::split(dcant$depth) %>% 
  lapply(raster::rasterFromXYZ) %>% 
  raster::brick() %>% 
  raster::setZ(z = unique(dcant$depth), name = "volume")

dcant_raster <- dcant %>% 
  select(lon, lat, dcant) %>%
  base::split(dcant$depth) %>% 
  lapply(raster::rasterFromXYZ) %>% 
  raster::brick() %>% 
  raster::setZ(z = unique(dcant$depth), name = "dcant")

# assign coordinate reference system
raster::crs(dcant_raster) <- coord_ref
raster::crs(volume_raster) <- coord_ref

# assign NA values
raster::NAvalue(dcant_raster) <- -9999
raster::NAvalue(dcant_raster)
raster::NAvalue(volume_raster) <- -9999
raster::NAvalue(volume_raster)

# check object
dim(dcant_raster)
raster::nbands(dcant_raster)
raster::nlayers(dcant_raster)
names(dcant_raster) #get the names of layers
raster::getZ(dcant_raster)

# write netcdf file
raster::writeRaster(
  dcant_raster,
  filename = paste0(path_preprocessing,
                    "dcant_Gruber2019_1994-2007_v20220608.nc"),
  overwrite = T
)

raster::writeRaster(
  volume_raster,
  filename = paste0(path_preprocessing,
                    "volume_Gruber2019_1994-2007_v20220608.nc"),
  overwrite = T
)


# modify created netcdf files
library(ncdf4)

# dcant file

# open file in writing mode
dcant_reopen <- nc_open(
  paste0(path_preprocessing,
         "dcant_Gruber2019_1994-2007_v20220608.nc"),
  write = TRUE)

dcant_reopen
print(dcant_reopen)
names(dcant_reopen$var)

# add units
ncatt_get(dcant_reopen, varid = "dcant")
ncatt_put(dcant_reopen, varid = "dcant",
          attname = "units", attval = "mol m-3")

ncatt_get(dcant_reopen, varid = "z")
ncatt_put(dcant_reopen, varid = "z",
          attname = "units", attval = "metres")

nc_close(dcant_reopen)


# volume file

# open file in writing mode
volume_reopen <- nc_open(
  paste0(path_preprocessing,
         "volume_Gruber2019_1994-2007_v20220608.nc"),
  write = TRUE)

volume_reopen
print(volume_reopen)
names(volume_reopen$var)

# add units
ncatt_get(volume_reopen, varid = "volume")
ncatt_put(volume_reopen, varid = "volume",
          attname = "units", attval = "m3")

ncatt_get(volume_reopen, varid = "z")
ncatt_put(volume_reopen, varid = "z",
          attname = "units", attval = "metres")

nc_close(volume_reopen)



# final check dcant

dcant_reopen <- tidync(
  paste0(path_preprocessing,
         "dcant_Gruber2019_1994-2007_v20220608.nc")) %>% 
  hyper_tibble()

dcant_reopen %>% 
  filter(z == 0) %>% 
  ggplot(aes(longitude, latitude, fill=dcant)) +
  geom_raster() +
  scale_fill_viridis_c()

dcant_reopen %>% 
  filter(longitude == 0.5) %>% 
  ggplot(aes(latitude, z, z=dcant)) +
  scale_y_reverse() +
  geom_contour_filled() +
  scale_fill_viridis_d()


dcant_reopen <- read_ncdf(
  paste0(path_preprocessing,
         "dcant_Gruber2019_1994-2007_v20220608.nc"))

plot(dcant_reopen,
     axes = TRUE)


# final check volume

volume_reopen <- tidync(
  paste0(path_preprocessing,
         "volume_Gruber2019_1994-2007_v20220608.nc")) %>% 
  hyper_tibble()

volume_reopen %>% 
  filter(z == 0) %>% 
  ggplot(aes(longitude, latitude, fill=volume)) +
  geom_raster() +
  scale_fill_viridis_c()

volume_reopen %>% 
  filter(longitude == 200.5) %>% 
  ggplot(aes(latitude, z, z=volume)) +
  scale_y_reverse() +
  geom_contour_filled() +
  scale_fill_viridis_d()


volume_reopen <- read_ncdf(
  paste0(path_preprocessing,
         "volume_Gruber2019_1994-2007_v20220608.nc"))

plot(volume_reopen,
     axes = TRUE)

```



