---
title: "Ocean regions"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r library, include=FALSE}
library(tidyverse)
library(tidync)
library(reticulate)
library(oce)
library(gsw)
library(geosphere)
library(patchwork)
library(stars)
library(khroma)
library(terra)
```

```{r define_paths, include=FALSE}

path_woa2018        <- "/nfs/kryo/work/updata/woa2018/"
path_reccap2 <-  "/nfs/kryo/work/updata/reccap2/"

path_root           <- "/nfs/kryo/work/jenmueller/emlr_cant"
path_functions      <- paste(path_root, "/utilities/functions/", sep = "")
path_files          <- paste(path_root, "/utilities/files/", sep = "")
path_preprocessing  <- paste(path_root, "/observations/preprocessing/", sep = "")

```

```{r read_params_global, include = FALSE}
params_global <-
  read_rds(paste(path_files,
                 "params_global.rds",
                 sep = ""))
```

```{r read_functions, include = FALSE}
source(paste(path_functions,
             "plotting_functions.R",
             sep = ""))
```

```{r set_ggplot_theme_global, include = FALSE}
theme_set(theme_bw())
```

# Data source
- Data source: [World Ocean Atlas 2018](https://www.nodc.noaa.gov/OC5/woa18/woa18data.html){target="_blank"}


```{r read_reccap2_region_mask, eval=FALSE}

region_masks_all <-
  read_ncdf(paste(path_reccap2, "RECCAP2_region_masks_all_v20210412.nc", sep = "")) %>%
  as_tibble()

# region_masks_all_seamask <- region_masks_all %>%
#   select(lat, lon, seamask)

region_masks_all <- region_masks_all %>%
  select(-seamask)

region_masks_all <- region_masks_all %>% 
  mutate(arctic = if_else(arctic != 0 & atlantic != 0, 0, arctic),
         southern = if_else(southern != 0 & atlantic != 0, 0, southern),
         southern = if_else(southern != 0 & pacific != 0, 0, southern),
         southern = if_else(southern != 0 & indian != 0, 0, southern))

region_masks_all <- region_masks_all %>%
  pivot_longer(open_ocean:southern,
               names_to = "region",
               values_to = "value") %>%
  mutate(value = as.factor(value)) %>% 
  mutate(lon = if_else(lon < 20, lon + 360, lon))

region_masks_all %>%
  filter(value != 0,
         region != "open_ocean") %>%
  ggplot(aes(lon, lat, fill = region)) +
  geom_raster() +
  scale_fill_brewer(palette = "Dark2") +
  coord_quickmap(expand = 0)

reccap2_region_mask <- region_masks_all %>%
  filter(value != 0,
         region != "open_ocean") %>% 
  select(lon, lat, region)

rm(region_masks_all)

```



# WOA Masks

## Land

### Read mask

The land sea mask with 1x1° resolution from the file `landsea_01.msk` was used.

```{r read_landsea_mask}

landsea_01 <- read_csv(
  paste(
    path_woa2018,
    "masks/landsea_01.msk",
    sep = ""),
  skip = 1,
  col_types = list(.default = "d"))

```

### Label

According to the [WOA18 documentation](https://data.nodc.noaa.gov/woa/WOA18/DOC/woa18documentation.pdf){target="_blank"} document:  

*"The landsea_XX.msk contains the standard depth level number at which the bottom of the ocean is first encountered at each quarter-degree or one-degree square for the entire world.  Land will have a value of 1, corresponding to the surface."*

The landmask was derived as coordinates with value 1.

```{r format_landmask}

landmask <- landsea_01 %>%
  mutate(region = if_else(Bottom_Standard_Level == "1",
                          "land", "ocean")) %>%
  select(-Bottom_Standard_Level)

landmask <- landmask %>%
  rename(lat = Latitude,
         lon = Longitude) %>%
  mutate(lon = if_else(lon < 20, lon + 360, lon)) %>%
  filter(lat >= params_global$lat_min,
         lat <= params_global$lat_max
         )

landseamask <- landmask

landmask <- landmask %>%
  filter(region == "land") %>%
  select(-region)

rm(landsea_01)

```

## Basins

### Read mask

The surface mask (0m) with 1x1° resolution from the file `basinmask_01.msk` was used.

```{r read_and_format_basinmask}

basinmask_01 <- read_csv(
  paste(
    path_woa2018,
    "masks/basinmask_01.msk",
    sep = ""),
  skip = 1,
  col_types = list(.default = "d"))

basinmask_01 <- basinmask_01 %>% 
  select(Latitude:Basin_0m) %>% 
  mutate(Basin_0m = as.factor(Basin_0m)) %>% 
  rename(lat = Latitude, lon = Longitude)
  
```


### Basins for budgets

According to [WOA FAQ](https://www.nodc.noaa.gov/OC5/WOD/wod-woa-faqs.html){target="_blank"} website and [WOA18 documentation](https://data.nodc.noaa.gov/woa/WOA18/DOC/woa18documentation.pdf){target="_blank"}, number codes in the mask files were used to assign ocean basins as follows:  

Atlantic Ocean:  

- 1: Atlantic Ocean
- 10: Southern Ocean between 63°W and 20°E
- 11: Arctic Ocean (restricted by northern latitude limit `r params_global$lat_max`N)

Indian Ocean:  

- 3: Indian Ocean 
- 10: Southern Ocean between 20°E and 147°E
- 56: Bay of Bengal

Pacific Ocean:  

- 2: Pacific Ocean
- 10: Southern Ocean between 147°E and 63°W
- 12: Sea of Japan; currently not included

```{r assign_standard_basin_labels}

# assign basin labels
basinmask_01 <- basinmask_01 %>%
  filter(Basin_0m %in% c("1", "2", "3", "10", "11", "12", "56")) %>%
  mutate(
    basin_AIP = "none",
    basin_AIP = case_when(
      Basin_0m == "1" ~ "Atlantic",
      Basin_0m == "10" & lon >= -63 & lon < 20 ~ "Atlantic",
      Basin_0m == "11" ~ "Atlantic",
      Basin_0m == "3" ~ "Indian",
      Basin_0m == "56" ~ "Indian",
      Basin_0m == "10" & lon >= 20 & lon < 147 ~ "Indian",
      Basin_0m == "2" ~ "Pacific",
      Basin_0m == "12" ~ "Pacific",
      Basin_0m == "10" &
        lon >= 147 | lon < -63 ~ "Pacific"
    )
  )

# apply northern latitude boundary
basinmask_01 <- basinmask_01 %>%
  filter(lat <= params_global$lat_max)

# harmonize lon scale
basinmask_01 <- basinmask_01  %>%
  mutate(lon = if_else(lon < 20, lon + 360, lon))

# prepare basin mask excluding Sea of Japa
basinmask_01_excl_Japan <- basinmask_01  %>%
  filter(Basin_0m != "12")


basinmask_01_excl_Japan <- basinmask_01_excl_Japan  %>%
  select(-Basin_0m)

basinmask_01 <- basinmask_01  %>%
  select(-Basin_0m)




```

```{r join_WOA_and_reccap2_mask, eval=FALSE}

basinmask_01 <- inner_join(basinmask_01,reccap2_region_mask)

basinmask_01 <- basinmask_01 %>% 
  mutate(basin_AIP = if_else(region == "arctic", "Arctic", basin_AIP))
  
```



```{r basin_AIP_masks_WOA18_map, fig.asp=0.5}

# generate base map, which is further used throughout the project
map <-
  ggplot() +
  geom_tile(data = landmask,
            aes(lon, lat), fill = "grey80") +
  coord_quickmap(expand = 0) +
  theme(axis.title = element_blank())

# plot basin_AIP map
map +
  geom_raster(data = basinmask_01,
              aes(lon, lat, fill = basin_AIP)) +
  scale_fill_brewer(palette = "Dark2")

# plot basin_AIP map
map +
  geom_raster(data = basinmask_01_excl_Japan,
              aes(lon, lat, fill = basin_AIP)) +
  scale_fill_brewer(palette = "Dark2")

```

```{r landsea_masks_WOA18_map, fig.asp=0.5}

# generate base map, which is further used throughout the project
ggplot() +
  geom_raster(data = landseamask,
              aes(lon, lat, fill = region)) +
  coord_quickmap(expand = 0) +
  scale_fill_brewer(palette = "Paired") +
  theme(axis.title = element_blank())

```

### Basins for MLR fitting

For the MLR fitting, ocean basins are further split up, as plotted below.

```{r assign_alternative_basin_labels}

# basinmask_01 <- basinmask_01 %>% 
#   select(-region)

# 4 basins incl arctic
# basinmask_04 <- basinmask_01 %>%
#   mutate(basin = basin_AIP) %>%
#   mutate(MLR_basins = "4")

# 1 basins
basinmask_01 <- basinmask_01 %>%
  # filter(basin_AIP != "Arctic") %>% 
  mutate(basin = "global",
         MLR_basins = "1")

# 2 basins
basinmask_2 <- basinmask_01 %>%
  mutate(basin = if_else(basin_AIP == "Atlantic",
                         "Atlantic",
                         "Indo-Pacific"),
         MLR_basins = "2")

# 5 basins
basinmask_5 <- basinmask_01 %>%
  mutate(
    basin = case_when(
      basin_AIP ==  "Atlantic" & lat > params_global$lat_equator ~ "N_Atlantic",
      basin_AIP ==  "Atlantic" & lat < params_global$lat_equator ~ "S_Atlantic",
      basin_AIP ==  "Pacific" & lat > params_global$lat_equator ~ "N_Pacific",
      basin_AIP ==  "Pacific" & lat < params_global$lat_equator ~ "S_Pacific",
      basin_AIP ==  "Indian" ~ "Indian"
    )
  ) %>%
  mutate(MLR_basins = "5")

# SO_2 basin separate
basinmask_SO_2 <- basinmask_01 %>%
  mutate(
    basin = if_else(basin_AIP == "Atlantic",
                         "Atlantic",
                         "Indo-Pacific"),
    basin = if_else(
      lat < params_global$lat_min_SO, "SO", basin)
  ) %>%
  mutate(MLR_basins = "SO_2")


# SO_5 basin separate
basinmask_SO_5 <- basinmask_01 %>%
  mutate(
    basin = case_when(
      basin_AIP ==  "Atlantic" & lat > 35 ~ "N_Atlantic",
      basin_AIP ==  "Atlantic" & lat < 35 & lat >= params_global$lat_min_SO ~ "Atlantic",
      basin_AIP ==  "Atlantic" & lat < params_global$lat_min_SO ~ "S_Atlantic",
      basin_AIP ==  "Pacific" & lat > 35 ~ "N_Pacific",
      basin_AIP ==  "Pacific" & lat < 35 & lat >= params_global$lat_min_SO ~ "Pacific",
      basin_AIP ==  "Pacific" & lat < params_global$lat_min_SO ~ "S_Pacific",
      basin_AIP ==  "Indian" & lat >= params_global$lat_min_SO ~ "Indian",
      basin_AIP ==  "Indian" & lat < params_global$lat_min_SO ~ "S_Indian"
    )) %>%
  mutate(MLR_basins = "SO_5")

# SO basin separate, with others being AIP
basinmask_SO_AIP <- basinmask_01 %>%
  mutate(
    basin = if_else(
      lat < params_global$lat_min_SO, "SO", basin_AIP)
  ) %>%
  mutate(MLR_basins = "SO_AIP")

# 3 basins
basinmask_AIP <- basinmask_01 %>%
  mutate(
    basin = basin_AIP) %>%
  mutate(MLR_basins = "AIP")


# join basin masks into one file
basinmask_all <- bind_rows(
  # basinmask_04,
  basinmask_01,
  basinmask_2,
  basinmask_5,
  basinmask_SO_2,
  basinmask_SO_5,
  basinmask_SO_AIP,
  basinmask_AIP
)

```

```{r basin_masks_WOA18_map, fig.asp=0.5}

for (i_MLR_basins in unique(basinmask_all$MLR_basins)) {
  # i_MLR_basins <- unique(basinmask_all$MLR_basins)[6]
  
  print(
    map +
      geom_raster(
        data = basinmask_all %>% filter(MLR_basins == i_MLR_basins),
        aes(lon, lat, fill = basin)
      ) +
      scale_fill_brewer(palette = "Dark2") +
      labs(title = paste("MLR basin label:", i_MLR_basins))
  )
  
}

```


## Global section

To plot sections from the North Atlantic south to the Southern Ocean, around Antarctica and back North across the Pacific Ocean, corresponding coordinates were subsetted from the basin mask and distances between coordinate grid points calculated.

### Line

```{r create_coordinates_of_section_global_line}

section <- basinmask_01 %>%
  select(lon, lat)

# subset individual section parts
Atl_NS <- section %>%
  filter(
    lon == params_global$lon_Atl_section,
    # lat <= params_global$lat_section_N,
    lat >= params_global$lat_section_S
  ) %>%
  arrange(-lat)

Atl_SO <- section %>%
  filter(lon > params_global$lon_Atl_section,
         lat == params_global$lat_section_S) %>%
  arrange(lon)

Pac_SO <- section %>%
  filter(lon < params_global$lon_Pac_section,
         lat == params_global$lat_section_S) %>%
  arrange(lon)

Pac_SN <- section %>%
  filter(
    lon == params_global$lon_Pac_section,
    # lat <= params_global$lat_section_N,
    lat >= params_global$lat_section_S
  ) %>%
  arrange(lat)

# join individual section parts
section_global_coordinates <- bind_rows(Atl_NS,
                     Atl_SO,
                     Pac_SO,
                     Pac_SN)

# convert to regular lon coordinates for distance calculation
section_global_coordinates <- section_global_coordinates %>%
  mutate(lon_180 = if_else(lon > 180, lon - 360, lon))

# calculate distance along section
section_global_coordinates <- section_global_coordinates %>%
  mutate(dist_int = distGeo(cbind(lon_180, lat)) / 1e6) %>%
  mutate(dist = cumsum(dist_int))

section_global_coordinates <- section_global_coordinates %>%
  select(lon, lat, dist) %>% 
  drop_na()

rm(Atl_NS, Atl_SO, Pac_SN, Pac_SO, section)

```

```{r section_global_map_line, fig.asp=0.5}

map +
  geom_point(data = section_global_coordinates,
             aes(lon, lat, col = dist)) +
  scale_colour_viridis_b(name = "Distance (Mm)")

```

### band

```{r create_coordinates_of_section_global_band}

section <- basinmask_01 %>%
  select(lon, lat)

# subset individual section parts
Atl_NS <- section %>%
  filter(
    lon >= params_global$lon_Atl_section - 5.5,
    lon <= params_global$lon_Atl_section + 4.5,
    # lat <= params_global$lat_section_N,
    lat >= params_global$lat_section_S
  ) %>%
  arrange(-lat)

Atl_SO <- section %>%
  filter(lon > params_global$lon_Atl_section,
         lat >= params_global$lat_section_S - 5.5,
         lat <= params_global$lat_section_S + 4.5) %>%
  arrange(lon)

Pac_SO <- section %>%
  filter(lon < params_global$lon_Pac_section,
         lat >= params_global$lat_section_S - 5.5,
         lat <= params_global$lat_section_S + 4.5) %>%
  arrange(lon)

Pac_SN <- section %>%
  filter(
    lon >= params_global$lon_Pac_section - 5.5,
    lon <= params_global$lon_Pac_section + 4.5,
    # lat <= params_global$lat_section_N,
    lat >= params_global$lat_section_S
  ) %>%
  arrange(lat)

# join individual section parts
section_global_coordinates_band <-
  bind_rows(Atl_NS %>% mutate(band = "Atlantic"),
            Atl_SO %>% mutate(band = "Southern"),
            Pac_SO %>% mutate(band = "Southern"),
            Pac_SN %>% mutate(band = "Pacific"))

section_global_coordinates_band <- 
  full_join(section_global_coordinates_band,
          section_global_coordinates)

rm(Atl_NS, Atl_SO, Pac_SN, Pac_SO, section)

```

```{r section_global_map_band, fig.asp=0.5}

map +
  geom_point(data = section_global_coordinates_band,
             aes(lon, lat, fill=band), alpha = 0.1, shape=21) +
  geom_point(data = section_global_coordinates,
             aes(lon, lat, col = dist)) +
  scale_colour_viridis_b(name = "Distance (Mm)")

```

### whole basin

```{r create_coordinates_of_section_global_basin}

section <- basinmask_01 %>%
  select(lon, lat, basin_AIP)

# subset individual section parts
Atl_NS <- section %>%
  filter(
    basin_AIP == "Atlantic",
    # lat <= params_global$lat_section_N,
    lat >= params_global$lat_section_S
  ) %>%
  arrange(-lat)

Atl_NS <- full_join(Atl_NS,
          section_global_coordinates %>% 
            filter(lon == params_global$lon_Atl_section))

Atl_NS <- Atl_NS %>%
  select(-basin_AIP) %>%
  mutate(band = "Atlantic")

SO <- basinmask_SO_AIP %>%
  filter((lon > params_global$lon_Atl_section |
         lon < params_global$lon_Pac_section) &
         basin == "SO",
         lat <= params_global$lat_section_S + 4.5,
         lat >= params_global$lat_section_S - 5.5)

SO <- left_join(
  SO,
  section_global_coordinates %>% 
    filter(lat == params_global$lat_section_S)
)

SO <- SO %>% 
  select(-c(basin_AIP:MLR_basins)) %>% 
  mutate(band = "Southern")

Pac_SN <- section %>%
  filter(
    basin_AIP == "Pacific",
    # lat <= params_global$lat_section_N,
    lat >= params_global$lat_section_S
  ) %>%
  arrange(lat)


Pac_SN <- full_join(Pac_SN,
          section_global_coordinates %>% 
            filter(lon == params_global$lon_Pac_section))

Pac_SN <- Pac_SN %>%
  select(-basin_AIP) %>%
  mutate(band = "Pacific")

# join individual section parts
section_global_coordinates_basin <-
  bind_rows(Atl_NS,
            SO,
            Pac_SN)


rm(Atl_NS, SO, Pac_SN, section)

```

```{r section_global_map_basin, fig.asp=0.5}

map +
  geom_point(data = section_global_coordinates_basin,
             aes(lon, lat, fill=band), alpha = 0.1, shape=21) +
  geom_point(data = section_global_coordinates,
             aes(lon, lat, col = dist)) +
  scale_colour_viridis_b(name = "Distance (Mm)")


```


## Write files

```{r write_files}

# land sea mask
landseamask %>%
  write_csv(paste(path_files,
                  "land_sea_mask_WOA18.csv",
                  sep = ""))

# basin mask
basinmask_all %>%
  write_csv(paste(path_files,
                  "basin_mask_WOA18.csv",
                  sep = ""))

# basin mask excluding Sea of Japan
basinmask_01_excl_Japan %>%
  write_csv(paste(path_files,
                  "basin_mask_WOA18_excl_Japan.csv",
                  sep = ""))

# global section
section_global_coordinates %>%
  write_csv(paste(path_files,
                  "section_global_coordinates.csv",
                  sep = ""))

# global section band
section_global_coordinates_band %>%
  write_csv(paste(path_files,
                  "section_global_coordinates_band.csv",
                  sep = ""))

# global section band
section_global_coordinates_basin %>%
  write_csv(paste(path_files,
                  "section_global_coordinates_basin.csv",
                  sep = ""))

# base map ggplot
map %>%
  write_rds(paste(path_files,
                  "map_landmask_WOA18.rds",
                  sep = ""))

```

