---
title: "World Ocean Atlas 2018"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r parent, child = "/nfs/kryo/work/jenmueller/emlr_cant/utilities/setup.Rmd"}
# this chunk runs the code stored in setup.Rmd
# if required, please refer to instructions given here:
# https://jdblischak.github.io/workflowr/articles/wflow-07-common-code.html
```

```{r library, include=FALSE}
library(tidync)
library(reticulate)
library(oce)
library(gsw)
library(geosphere)
library(stars)
```

```{r define_paths, include=FALSE}

path_woa2018        <- "/nfs/kryo/work/updata/woa2018/"
path_preprocessing  <- paste(path_root, "/observations/preprocessing/", sep = "")

```

# Data source
- Data source: [World Ocean Atlas 2018](https://www.nodc.noaa.gov/OC5/woa18/woa18data.html){target="_blank"}

# Climatology S and T

Copied from the WOA FAQ website, the file naming conventions is:  

PREF_DDDD_VTTFFGG.EXT, where:

- PREF: prefix
- DDDD: decade
- V: variable
- TT: time period
- FF: field type
- GG: grid (5deg- 5°, 01- 1°, 04 - 1/4°)
- EXT: file extention

Short description of two statistical fields in WOA

- Objectively analyzed climatologies are the objectively interpolated mean fields for oceanographic variables at standard - depth levels for the World Ocean.
- The statistical mean is the average of all unflagged interpolated values at each standard depth level for each variable - in each 1° square which contains at least one measurement for the given oceanographic variable.

Here, we use  

- Fields: objectively analyzed mean
- Decades: all decades
- Grid: 1 deg resolution

According to the [WOA18 documentation](https://data.nodc.noaa.gov/woa/WOA18/DOC/woa18documentation.pdf){target="_blank"} document:

*What are the units for temperature and salinity in the WOA18?*

*In situ temperatures used for WOA18 are not converted from their original scale, so there is a mix of IPTS-48, IPTS-68, and ITS-90 (and pre IPTS-48 temperatures). The differences between scales are small (on the order of 0.01°C) and should not have much effect on the climatological means, except, possibly at very deep depths. Values for salinity are on the Practical salinity scale (PSS-78). Pre-1978 salinity values converted from conductivity may have used a different salinity scale. Pre-conductivity salinities use the Knudsen method.* 


## Read nc files

```{r read_WOA18_salinity_temperature}

# temperature

WOA18_temp <- tidync(paste(
  path_woa2018,
  "temperature/decav/1.00/woa18_decav_t00_01.nc",
  sep = ""
))

WOA18_temp_tibble <- WOA18_temp %>%
  hyper_tibble()

WOA18_temp_tibble <- WOA18_temp_tibble  %>%
  select(temp = t_an, lon, lat, depth) %>%
  drop_na() %>%
  mutate(lon = if_else(lon < 20, lon + 360, lon))

# salinity

WOA18_sal <- tidync(paste(
  path_woa2018,
  "salinity/decav/1.00/woa18_decav_s00_01.nc",
  sep = ""
))

WOA18_sal_tibble <- WOA18_sal %>% hyper_tibble()

WOA18_sal_tibble <- WOA18_sal_tibble  %>%
  select(sal = s_an, lon, lat, depth) %>%
  drop_na() %>%
  mutate(lon = if_else(lon < 20, lon + 360, lon))

rm(WOA18_sal, WOA18_temp)

```

## Join predictors

```{r join_predictors}

WOA18_sal_temp <- full_join(WOA18_sal_tibble, WOA18_temp_tibble)
rm(WOA18_sal_tibble, WOA18_temp_tibble)

```

## Apply basin mask

```{r apply_basin_mask}

# use only three basin to assign general basin mask
# ie this is not specific to the MLR fitting

basinmask <- basinmask %>% 
  filter(MLR_basins == "2") %>% 
  select(lat, lon, basin_AIP)

# restrict predictor fields to basin mask grid

WOA18_sal_temp <- inner_join(WOA18_sal_temp, basinmask)

```

## Subset depth levels

```{r subset_depth_levels}

WOA18_sal_temp <- WOA18_sal_temp %>% 
  filter(depth %in% params_global$depth_levels_33)

```


## Potential temperature

Potential temperature is calculated as in input variable for the neutral density calculation.

### Calculation

```{r calculate_potential_temperature}

WOA18_sal_temp <- WOA18_sal_temp %>% 
  mutate(THETA = swTheta(salinity = sal,
                         temperature = temp,
                         pressure = depth,
                         referencePressure = 0,
                         longitude = lon - 180,
                         latitude = lat))

```


### Profile

Example profile from North Atlantic Ocean.

```{r WOA18_potential_temperature_profile}

WOA18_sal_temp %>%
  filter(lat == params_global$lat_Atl_profile,
         lon == params_global$lon_Atl_section) %>%
  ggplot() +
  geom_line(aes(temp, depth, col = "insitu")) +
  geom_point(aes(temp, depth, col = "insitu")) +
  geom_line(aes(THETA, depth, col = "theta")) +
  geom_point(aes(THETA, depth, col = "theta")) +
  scale_y_reverse() +
  scale_color_brewer(palette = "Dark2", name = "Scale")

```

### Section

```{r WOA18_pot_temperature_section}

p_section_global(
  df = WOA18_sal_temp,
  var = "THETA")

```


## Neutral density

Neutral density gamma was calculated with a Python script provided by Serazin et al (2011), which performs a polynomial approximation of the original gamma calculation.

### Calculation

```{r calculate_neutral_density}

# calculate pressure from depth

WOA18_sal_temp <- WOA18_sal_temp %>%
  mutate(CTDPRS = gsw_p_from_z(-depth,
                               lat))

# rename variables according to python script

WOA18_sal_temp_gamma_prep <- WOA18_sal_temp %>%
  rename(LATITUDE = lat,
         LONGITUDE = lon,
         SALNTY = sal)

# load python scripts

source_python(paste(
  path_functions,
  "python_scripts/Gamma_GLODAP_python.py",
  sep = ""
))

# calculate gamma

WOA18_sal_temp_gamma_calc <-
  calculate_gamma(WOA18_sal_temp_gamma_prep)

# reverse variable naming

WOA18_sal_temp <- WOA18_sal_temp_gamma_calc %>%
  select(-c(CTDPRS, THETA)) %>%
  rename(
    lat = LATITUDE,
    lon = LONGITUDE,
    sal = SALNTY,
    gamma  = GAMMA
  )

WOA18_sal_temp <- as_tibble(WOA18_sal_temp)

rm(WOA18_sal_temp_gamma_calc, WOA18_sal_temp_gamma_prep)

```

```{r calculate_sigma_density}

# calculate pressure from depth

WOA18_sal_temp_dens <- WOA18_sal_temp %>%
  mutate(
    p = gsw_p_from_z(z = -depth, latitude = lat),
    CT = gsw_CT_from_t(SA = sal, t = temp, p = p),
    sigma0 = gsw_sigma0(SA = sal, CT = CT),
    sigma1 = gsw_sigma1(SA = sal, CT = CT),
    sigma4 = gsw_sigma4(SA = sal, CT = CT),
    rho = gsw_rho(SA = sal, CT = CT, p = p) -1000
    )

WOA18_sal_temp_dens <- WOA18_sal_temp_dens %>% 
  select(lon, lat, depth, basin_AIP,
         gamma, starts_with("sigma"), rho) %>% 
  pivot_longer(gamma:rho,
               names_to = "estimate",
               values_to = "value")

WOA18_sal_temp_dens %>%
  group_split(estimate) %>%
  # head(1) %>%
  map(~ p_map_climatology(df = .x,
                          var = "value",
                          title_text = unique(.x$estimate)))

WOA18_sal_temp_dens %>%
  group_split(estimate) %>%
  # head(1) %>%
  map(~ p_section_global(df = .x,
                          var = "value",
                          title_text = unique(.x$estimate)))

WOA18_sal_temp_dens <- WOA18_sal_temp_dens %>% 
  arrange(estimate) %>% 
  group_by(lon, lat, depth) %>% 
  mutate(delta_value = value - first(value),
         delta_estimate = paste(estimate, first(estimate), sep = "-")) %>% 
  ungroup()




WOA18_sal_temp_dens %>%
  group_split(delta_estimate) %>%
  # head(2) %>%
  map( ~ p_map_climatology(
    df = .x,
    var = "delta_value",
    col = "divergent",
    title_text = unique(.x$delta_estimate)
  ))

WOA18_sal_temp_dens %>%
  group_split(delta_estimate) %>%
  # head(2) %>%
  map( ~ p_section_global(
    df = .x,
    var = "delta_value",
    col = "divergent",
    subtitle_text = unique(.x$delta_estimate)
  ))

rm(WOA18_sal_temp_dens)

```

## Write file

```{r write_WOA18_sal_temp_predictor_file}

WOA18_sal_temp %>%
  write_csv(paste(path_preprocessing,
                  "WOA18_sal_temp.csv",
                  sep = ""))

```

## Temperature plots

Below, following subsets of the climatologies are plotted for all relevant parameters:  

- Horizontal planes at `r params_global$depth_levels`m
- Global section as defined above and indicated as white lines in maps.

### Surface map

```{r temperature_surface_WOA18_map, fig.asp=0.6}

p_map_climatology(
  df = WOA18_sal_temp,
  var = "temp")

```

### Section

```{r temperature_sections_WOA18}

p_section_global(
  df = WOA18_sal_temp,
  var = "temp")

```


## Salinity plots

### Surface map

```{r salinity_surface_WOA18_map}

p_map_climatology(
  df = WOA18_sal_temp,
  var = "sal")

```

### Section

```{r salinity_sections_WOA18}

p_section_global(
  df = WOA18_sal_temp,
  var = "sal")

```


## Neutral density plots

### Surface map

```{r neutral_density_surface_WOA18_map}

p_map_climatology(
  df = WOA18_sal_temp,
  var = "gamma")

```

### Section

```{r neutral_density_sections_WOA18}

p_section_global(
  df = WOA18_sal_temp,
  var = "gamma")

```


# Climatology Nuts and O2

## Read nc files

Data are read-in looping over all relevant files, thereby reproducing the same subsetting steps as applied above to the salintity and temperature fields.

```{r read_WOA18_nuts_O2_files}

# Keep grid cells of WOA18 sal temp data set, to join with
WOA18_nuts_O2 <-
  WOA18_sal_temp %>%
  select(lon, lat, depth)
rm(WOA18_sal_temp)

# create file list
file_list <- c(
  paste(path_woa2018, "phosphate/all/1.00/woa18_all_p00_01.nc", sep = ""),
  paste(path_woa2018, "nitrate/all/1.00/woa18_all_n00_01.nc", sep = ""),
  paste(path_woa2018, "silicate/all/1.00/woa18_all_i00_01.nc", sep = ""),
  paste(path_woa2018, "oxygen/all/1.00/woa18_all_o00_01.nc", sep = ""),
  paste(path_woa2018, "AOU/all/1.00/woa18_all_A00_01.nc", sep = "")
)

# read, plot and join data sets while looping over file list
for (file in file_list) {
  # file <- file_list[1]

  # open file
  WOA18 <- tidync(file)
  WOA18_tibble <- WOA18 %>% hyper_tibble()
  
  # extract parameter name
  parameter <- str_split(file, pattern = "00_", simplify = TRUE)[1]
  parameter <- str_split(parameter, pattern = "all_", simplify = TRUE)[2]
  parameter <- paste(parameter, "_an", sep = "")
  print(file)
  
  WOA18_tibble <- WOA18_tibble  %>%
    select(all_of(parameter),
           lon, lat, depth) %>%
    mutate(lon = if_else(lon < 20, lon + 360, lon))
  
  # apply general basin mask
  WOA18_tibble <- inner_join(WOA18_tibble, basinmask)
  
  # subset depth levels
  WOA18_tibble <- WOA18_tibble %>%
    filter(depth %in% params_global$depth_levels_33)
  
  
  # join with previous WOA data and keep only rows in existing data frame
  # this is equal to applying the basinmask
  WOA18_nuts_O2 <- left_join(
    x = WOA18_nuts_O2,
    y = WOA18_tibble)

  # plot maps
  print(
    p_map_climatology(
      df = WOA18_nuts_O2,
      var = parameter)
    )
  
  # plot sections
  print(p_section_global(
    df = WOA18_nuts_O2,
    var = parameter
  ))
  
}


```


## Write file

```{r write_WOA18_nuts_O2_predictor_file}

WOA18_nuts_O2 %>%
  rename(phosphate = p_an,
         nitrate = n_an,
         silicate = i_an,
         oxygen = o_an,
         aou = A_an) %>% 
  write_csv(paste(path_preprocessing,
                  "WOA18_nuts_O2.csv",
                  sep = ""))

```
