---
title: "Anthropogenic CO2 from 1994 to 2007"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r parent, child = "/nfs/kryo/work/jenmueller/emlr_cant/utilities/setup.Rmd"}
# this chunk runs the code stored in setup.Rmd
# if required, please refer to instructions given here:
# https://jdblischak.github.io/workflowr/articles/wflow-07-common-code.html
```

```{r define_paths}
path_sabine_2004    <- "/nfs/kryo/work/updata/glodapv1_1/GLODAP_gridded.data/"
path_preprocessing  <- paste(path_root, "/observations/preprocessing/", sep = "")
```

```{r load_libraries}
library(marelac)

```


# Data source

- Anthropogenic CO2 estimates (1800-1994) by Sabine et al. (2004) downloaded in August 2020 from [NOAA/NCEI Ocean Carbon Data System (OCADS)](ftp://ftp.nodc.noaa.gov/pub/data.nodc/ncei/ocads/data/0001644/){target="_blank"}

# Read nc files

```{r read_Sabine_2004}

# read text files
AnthCO2_data <-
  read_csv(
    paste(path_sabine_2004,
          "AnthCO2.data/AnthCO2.data.txt",
          sep = ""),
    col_names = FALSE,
    na = "-999",
    col_types = list(.default = "d")
  )

# read respective depth layers and convert to vector
Depth_centers <-
  read_file(paste(path_sabine_2004,
                  "Depth.centers.txt",
                  sep = ""))

Depth_centers <- Depth_centers %>%
  str_split(",") %>%
  as_vector()

# read respective latitudes and convert to vector
Lat_centers <-
  read_file(paste(path_sabine_2004, "Lat.centers.txt",
                  sep = ""))

Lat_centers <- Lat_centers %>%
  str_split(",") %>%
  as_vector()

# read respective longitudes and convert to vector
Long_centers <-
  read_file(paste(path_sabine_2004, "Long.centers.txt",
                  sep = ""))

Long_centers <- Long_centers %>%
  str_split(",") %>%
  as_vector()

# match lon, lat and depth vectors with Cant value file
names(AnthCO2_data) <- Lat_centers

Long_Depth <-
  expand_grid(depth = Depth_centers, lon = Long_centers) %>%
  mutate(lon = as.numeric(lon),
         depth = as.numeric(depth))

tcant_3d <- bind_cols(AnthCO2_data, Long_Depth)

# adjust file dimensions
tcant_3d <- tcant_3d %>%
  pivot_longer(1:180, names_to = "lat", values_to = "tcant") %>%
  mutate(lat = as.numeric(lat))

tcant_3d <- tcant_3d %>%
  drop_na()

# harmonize coordinates
tcant_3d <- tcant_3d %>%
  mutate(lon = if_else(lon < 20, lon + 360, lon))

rm(AnthCO2_data,
   Long_Depth,
   Depth_centers,
   Lat_centers,
   Long_centers)

```

# Apply basin mask

```{r apply_basin_mask}

# use only three basin to assign general basin mask
# ie this is not specific to the MLR fitting
basinmask <- basinmask %>% 
  filter(MLR_basins == "2") %>% 
  select(lat, lon, basin_AIP)

tcant_3d_unmasked <- tcant_3d
tcant_3d <- inner_join(tcant_3d, basinmask)


ggplot() +
  geom_tile(data = tcant_3d_unmasked %>% 
              distinct(lon, lat),
            aes(lon, lat, fill = "basin mask not applied")) +
  geom_tile(data = tcant_3d %>% 
              distinct(lon, lat),
            aes(lon, lat, fill = "basin mask applied")) +
  coord_quickmap()

rm(tcant_3d_unmasked)

```


# Calculation

## Column inventory

```{r calculate_inventory}

tcant_3d <- tcant_3d %>% 
  mutate(tcant_pos = if_else(tcant <= 0, 0, tcant))

tcant_inv_layers <- m_tcant_inv(tcant_3d)

tcant_inv <- tcant_inv_layers %>% 
  filter(inv_depth == params_global$inventory_depth_standard)

```

## Zonal mean section

```{r calculate_zonal_mean_section}

tcant_zonal <- m_zonal_mean_sd(tcant_3d)

```

```{r calculate_budgets}

m_dcant_budget(
  tcant_inv_layers %>%
    rename(dcant = tcant,
           dcant_pos = tcant_pos) %>%
    mutate(method = "total",
           data_source = "obs")) %>%
  select(-c(data_source, method)) %>% 
  group_by(estimate) %>%
  mutate(ratio = round(value / lag(value),3)) %>%
  ungroup() %>% 
  arrange(estimate, inv_depth)

```



# Plots


## Inventory map

```{r cant_pos_inventory_maps, fig.asp=0.6}

p_map_cant_inv(
  df = tcant_inv,
  var = "tcant_pos",
  breaks = seq(0,max(tcant_inv$tcant_pos),5))

p_map_cant_inv(
  df = tcant_inv,
  var = "tcant",
  breaks = seq(0,max(tcant_inv$tcant_pos),5))

```

## Horizontal plane maps

```{r cant_maps, fig.asp=0.6}

p_map_climatology(
  df = tcant_3d,
  var = "tcant",
  col = "divergent")

```

## Global section

```{r cant_sections}

p_section_global(
  df = tcant_3d,
  var = "tcant",
  col = "divergent")

```

## Sections at regular longitudes

```{r cant_sections_regular, fig.asp=2}

p_section_climatology_regular(
  df = tcant_3d,
  var = "tcant",
  col = "divergent")

```

## Write files

```{r write_Sabine_2004_cant_file}

tcant_3d %>%
  write_csv(paste(path_preprocessing,
                  "S04_tcant_3d.csv", sep = ""))

tcant_inv %>%
  write_csv(paste(path_preprocessing,
                  "S04_tcant_inv.csv", sep = ""))

tcant_zonal %>%
  write_csv(paste(path_preprocessing,
                  "S04_tcant_zonal.csv", sep = ""))


```


# Anomalous changes

```{r read_G19_cant_file}

tcant_inv_S04 <- tcant_inv

dcant_inv_G19 <- read_csv(paste(path_preprocessing,
                               "G19_dcant_inv.csv", sep = ""))

```

# Comparison of previous estimates

Cant inventory estimates of S04 (Sabine et al, 2004) and G19 (Gruber et al, 2019) were compared.

## Merge data sets

```{r merge_cant_data_sets}

cant_inv <- full_join(dcant_inv_G19 %>%
                        mutate(estimate = "G19") %>% 
                        rename(cant_pos = dcant_pos) %>% 
                        select(-dcant),
                      tcant_inv_S04 %>% 
                        mutate(estimate = "S04") %>% 
                        rename(cant_pos = tcant_pos) %>% 
                        select(-tcant))

rm(dcant_inv_G19, tcant_inv_S04)

```

## Inventory maps

Spanning different time periods, the Cant inventories differ in magnitude. Please note, that we refer to cant_pos here, but strictly speaking we compare dcant and tcant.

```{r tcant_inv_comparison, fig.asp=1}
 
map +
  geom_raster(data = cant_inv,
              aes(lon, lat, fill = cant_pos)) +
  scale_fill_viridis_c() +
  facet_wrap( ~ estimate, ncol = 1) +
  theme(
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank()
  )

```

## C~ant~ budgets

Global Cant inventories were estimated in Pg-C. Please note that here we only added positive Cant values in the upper `r params_global$inventory_depth`m and do not apply additional corrections for areas not covered.

```{r calculate_global_inventory}

cant_inv <- cant_inv %>% 
  mutate(surface_area = earth_surf(lat, lon),
         cant_pos_grid = cant_pos*surface_area)

cant_inv_budget <- cant_inv %>% 
  group_by(estimate, basin_AIP) %>% 
  summarise(cant_pos_total = sum(cant_pos_grid)*12*1e-15,
            cant_pos_total = round(cant_pos_total,1)) %>% 
  ungroup() %>% 
  pivot_wider(values_from = cant_pos_total, names_from = basin_AIP) %>% 
  mutate(total = Atlantic + Indian + Pacific)

cant_inv_budget

```


## Relative inventories

```{r calculate_tcant_inv_ratio}

cant_inv_wide <- cant_inv %>%
  pivot_wider(values_from = c(cant_pos, cant_pos_grid),
              names_from = estimate)

cant_inv_wide <- cant_inv_wide %>% 
  drop_na() %>% 
  mutate(G19_rel = cant_pos_grid_G19 / sum(cant_pos_grid_G19),
         S04_rel = cant_pos_grid_S04 / sum(cant_pos_grid_S04),
         cant_ratio_rel = G19_rel / S04_rel)

cant_inv_rel <- cant_inv_wide %>%
  pivot_longer(
    cols = c(G19_rel, S04_rel),
    names_to = "estimate",
    values_to = "cant_pos_rel"
  )

```



```{r tcant_inv_rel_maps, fig.asp=1}

map +
  geom_raster(data = cant_inv_rel,
              aes(lon, lat, fill = cant_pos_rel*100)) +
  scale_fill_viridis_c() +
  facet_wrap( ~ estimate, ncol = 1) +
  theme(
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank()
  )

```

## Relative inventory ratios

```{r tcant_inv_ratio_map, fig.asp=0.6}

map +
  geom_contour_filled(data = cant_inv_wide %>%
  filter(cant_ratio_rel < 10,
         cant_ratio_rel > 0.1),
                      aes(lon, lat, z = log10(cant_ratio_rel))) +
  coord_quickmap(expand = 0) +
  scale_fill_brewer(palette = "RdBu", direction = -1) +
  labs(title = "Cant inventory distribution | 1994-2007 vs preind-1994",
       subtitle = "Log ratio of relative contributions to total inventory") +
  theme(
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    legend.title = element_blank()
  )

```


