---
title: "Anthropogenic CO2 from 1994 to 2007"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r parent, child = "/nfs/kryo/work/jenmueller/emlr_cant/utilities/setup_obs.Rmd"}
# this chunk runs the code stored in setup.Rmd
# if required, please refer to instructions given here:
# https://jdblischak.github.io/workflowr/articles/wflow-07-common-code.html
```

```{r define_paths, include=FALSE}
path_gruber_2019    <- "/nfs/kryo/work/updata/cant_gruber_2019/"
path_preprocessing  <- paste(path_root, "/observations/preprocessing/", sep = "")
```

```{r load_libraries_specific, include=FALSE}
library(tidync)
library(stars)
```


# Data source

- Anthropogenic CO2 estimates (1994-2007) by Gruber et al. (2019) downloaded in August 2020 from [NOAA/NCEI Ocean Carbon Data System (OCADS)](https://www.nodc.noaa.gov/ocads/oceans/ndp_100/ndp100.html){target="_blank"}

# Read nc files

Here, we use the standard case V101 for public and raw data sets.

## Public data sets

The publicly available data sets contain only positive Cant estimates.

### 3d fields

```{r read_Gruber_dcant}

# open file
dcant <- tidync(paste(
  path_gruber_2019,
  "dcant_emlr_cstar_gruber_94-07_vs1.nc",
  sep = ""
))

# read gamma field as tibble
dcant <- dcant %>%  activate(GAMMA_DENS)
dcant_gamma <- dcant %>% hyper_tibble()

# read delta cant field
dcant <- dcant %>%  activate(DCANT_01)
dcant <- dcant %>% hyper_tibble()

# join cant and gamma fields
dcant <- left_join(dcant, dcant_gamma)

# harmonize column names and coordinates
dcant <- dcant %>% 
  rename(lon = LONGITUDE,
         lat = LATITUDE,
         depth = DEPTH,
         gamma = GAMMA_DENS,
         dcant_pos = DCANT_01) %>% 
  mutate(lon = if_else(lon < 20, lon + 360, lon))

rm(dcant_gamma)

```

### Column inventories

```{r read_Gruber_dcant_inv_v101}

dcant_inv <- tidync(paste(
  path_gruber_2019,
  "inv_dcant_emlr_cstar_gruber_94-07_vs1.nc",
  sep = ""
))

dcant_inv <- dcant_inv %>%  activate(DCANT_INV01)
dcant_inv <- dcant_inv %>% hyper_tibble()

# harmonize column names and coordinates
dcant_inv <- dcant_inv %>% 
  rename(lon = LONGITUDE,
         lat = LATITUDE,
         dcant_pos = DCANT_INV01) %>% 
  mutate(lon = if_else(lon < 20, lon + 360, lon))

```

```{r read_Gruber_dcant_inv_all}

dcant_inv_all <- read_ncdf(
  paste(
    path_gruber_2019,
    "inv_dcant_emlr_cstar_gruber_94-07_vs1.nc",
    sep = ""
  ),
  var = sprintf("DCANT_INV%02d", seq(1, 14, 1)),
  make_units = FALSE
)

dcant_inv_all <- dcant_inv_all %>% as_tibble()

dcant_inv_all <- dcant_inv_all %>% 
  pivot_longer(DCANT_INV01:DCANT_INV14,
               names_to = "Version_ID",
               values_to = "dcant_pos",
               names_prefix = "DCANT_INV")

# harmonize column names and coordinates
dcant_inv_all <- dcant_inv_all %>%
  rename(lon = LONGITUDE,
         lat = LATITUDE) %>%
  mutate(lon = if_else(lon < 20, lon + 360, lon))

```

## Raw data

Internally available data sets also contain negative Cant estimates, as they are generated in the "raw" output of the eMLR mapping step.

```{r read_Gruber_V101}

# open v 101 file
V101 <- tidync(paste(path_gruber_2019,
                     "Cant_V101new.nc",
                     sep = ""))

# create tibble
V101 <- V101 %>%  activate(Cant)
V101 <- V101 %>% hyper_tibble()

# harmonize column names and coordinates
V101 <- V101 %>% 
  rename(lon = longitude,
         lat = latitude,
         dcant = Cant) %>% 
  filter(dcant != -999) %>% 
  mutate(lon = if_else(lon < 20, lon + 360, lon))

```


# Apply basin mask

```{r apply_basin_mask}

# use only three basin to assign general basin mask
# ie this is not specific to the MLR fitting
basinmask <- basinmask %>% 
  filter(MLR_basins == "2") %>% 
  select(lat, lon, basin_AIP)

dcant <- inner_join(dcant, basinmask)
dcant_inv_publ <- inner_join(dcant_inv, basinmask)
dcant_inv_all <- inner_join(dcant_inv_all, basinmask)
V101 <- inner_join(V101, basinmask)

```

# Join pos and all delta Cant

```{r join_V101_and_all}

# join files
dcant_3d <- inner_join(dcant, V101)

rm(dcant, V101)

```

# Zonal mean section

```{r calculate_zonal_mean_section}

dcant_zonal <- m_zonal_mean_sd(dcant_3d)

```

# Column inventory 

## Calculation

```{r calculate_column_inventory}

dcant_inv_layers <- m_dcant_inv(dcant_3d)

dcant_inv <- dcant_inv_layers %>% 
  filter(inv_depth == params_global$inventory_depth_standard)

```

## Plots

### All Cant

```{r G19_cant_inv_maps, fig.asp=0.6}

p_map_cant_inv(
  df = dcant_inv,
  var = "dcant",
  col = "divergent")

```

### Pos Cant

```{r G19_cant_pos_inventory_maps, fig.asp=0.6}

p_map_cant_inv(
  df = dcant_inv,
  var = "dcant_pos")

```

```{r G19_cant_pos_inventory_maps_all, fig.asp=1.5}

p_map_cant_inv(
  df = dcant_inv_all %>% mutate(dcant_pos = dcant_pos*(10/13)),
  var = "dcant_pos") +
  facet_wrap(~ Version_ID, ncol = 2)

```

### Published inventories

```{r G19_cant_pos_inventory_maps_published, fig.asp=0.6}

p_map_cant_inv(
  df = dcant_inv,
  var = "dcant_pos")

```

### Published vs calculated

```{r G19_cant_offset_inventory, fig.asp=0.6}

# join published and calculated data sets
dcant_inv_offset <- inner_join(
  dcant_inv %>% rename(dcant_re = dcant_pos),
  dcant_inv_publ %>% rename(dcant_pub = dcant_pos)
)

# calculate offset
dcant_inv_offset <- dcant_inv_offset %>% 
  mutate(dcant_offset = dcant_re - dcant_pub)

# plot map
p_map_cant_inv(
  df = dcant_inv_offset,
  var = "dcant_offset",
  col = "divergent",
  breaks = seq(-3, 3, 0.25)
)

rm(dcant_inv_offset, dcant_inv_publ)

```


# Horizontal plane maps

## All Cant

```{r Cant_maps, fig.asp=0.6}

p_map_climatology(
  df = dcant_3d,
  var = "dcant",
  col = "divergent")

```

## Positive Cant

```{r Cant_maps_V101, fig.asp=0.6}

p_map_climatology(
  df = dcant_3d,
  var = "dcant_pos")

```

## Neutral density

```{r Gamma_maps_V101, fig.asp=0.6}

p_map_climatology(
  df = dcant_3d,
  var = "gamma")

```


# Zonal mean section plot

## Positive Cant

```{r Cant_pos_zonal_mean_sections}

dcant_zonal %>%
  group_split(basin_AIP) %>%
  head(1) %>%
  map(
    ~ p_section_zonal(
      df = .x,
      var = "dcant_pos_mean",
      plot_slabs = "n",
      subtitle_text = paste("Basin:", unique(.x$basin_AIP))
    )
  )

```

# Global sections plot

## All Cant

```{r Cant_sections_V101}

p_section_global(
  df = dcant_3d,
  var = "dcant",
  col = "divergent")

```

## Positive Cant

```{r Cant_sections}

p_section_global(
  df = dcant_3d,
  var = "dcant_pos")

```


# Sections at regular longitudes

## All Cant

```{r Cant_sections_regular, fig.asp=1.7}

p_section_climatology_regular(
  df = dcant_3d,
  var = "dcant",
  col = "divergent")

```

## Positive Cant

```{r Cant_sections_regular_V101, fig.asp=1.7}

p_section_climatology_regular(
  df = dcant_3d,
  var = "dcant_pos")

```

## Neutral density

```{r Gamma_sections_regular_V101, fig.asp=1.7}

p_section_climatology_regular(
  df = dcant_3d,
  var = "gamma")

```

# Write files

```{r write_Gruber_Cant_file}

dcant_3d %>%
  write_csv(paste(path_preprocessing,
                  "G19_dcant_3d.csv",
                  sep = ""))

dcant_inv %>%
  write_csv(paste(path_preprocessing,
                  "G19_dcant_inv.csv",
                  sep = ""))

dcant_inv_all %>%
  write_csv(paste(path_preprocessing,
                  "G19_dcant_inv_all.csv",
                  sep = ""))

dcant_zonal %>%
  write_csv(paste(path_preprocessing,
                  "G19_dcant_zonal.csv",
                  sep = ""))

```


# RECCAP2-ocean


```{r write_Gruber_dcant_example_for_Michael_Summer, eval=FALSE}

# https://pjbartlein.github.io/REarthSysSci/netCDF.html#create-and-write-a-netcdf-file

# open file
dcant <- tidync(paste(
  path_gruber_2019,
  "dcant_emlr_cstar_gruber_94-07_vs1.nc",
  sep = ""
))

# read delta cant field
dcant <- dcant %>%  activate(DCANT_01)
dcant <- dcant %>% hyper_tibble(na.rm = FALSE)

# harmonize column names and coordinates
dcant <- dcant %>% 
  rename(lon = LONGITUDE,
         lat = LATITUDE,
         depth = DEPTH,
         dcant = DCANT_01) %>% 
  arrange(lon, lat, depth) %>% 
  relocate(dcant, .after = last_col())

dcant %>% 
  filter(depth == 0,
         lon > 125,
         lon < 175,
         lat < -25) %>% 
  ggplot(aes(lon, lat, fill=dcant)) +
  geom_raster() +
  scale_fill_viridis_c(na.value = "transparent")

unique(dcant$depth)

dcant %>%
  filter(lon > 125,
         lon < 175,
         lat < -25) %>% 
  write_csv(paste(path_preprocessing,
                  "G19_dcant_3d_example.csv",
                  sep = ""))

```


```{r write_Gruber_dcant_netcdf, eval=FALSE}

# https://pjbartlein.github.io/REarthSysSci/netCDF.html#create-and-write-a-netcdf-file

# open file
dcant <- tidync(paste(
  path_gruber_2019,
  "dcant_emlr_cstar_gruber_94-07_vs1.nc",
  sep = ""
))

# read delta cant field
dcant <- dcant %>%  activate(DCANT_01)
dcant <- dcant %>% hyper_tibble(na.rm = FALSE)

# harmonize column names and coordinates
dcant <- dcant %>% 
  rename(lon = LONGITUDE,
         lat = LATITUDE,
         depth = DEPTH,
         dcant_pos = DCANT_01) %>% 
  arrange(lon, lat, depth) %>% 
  relocate(dcant_pos, .after = last_col())

lon <- unique(dcant$lon)
lat <- unique(dcant$lat)
depth <- unique(dcant$depth)

nlon <- length(lon)
nlat <- length(lat)
ndepth <- length(depth)

dcant_mat <- as.matrix(dcant[4])
dcant_array <- array(dcant_mat, dim=c(nlon, nlat, ndepth))
dim(dcant_array)

library(lattice)
library(RColorBrewer)
grid <- expand.grid(lon=lon, lat=lat)

levelplot(dcant_array[,,1] ~ lon * lat, data=grid)
image(lon,lat,dcant_array[,,1], col=rev(brewer.pal(10,"RdBu")))

levelplot(dcant_array[,,1] ~ lon * lat, data=grid, at=cutpts, cuts=11, pretty=T, 
  col.regions=(rev(brewer.pal(10,"RdBu"))), main="Mean July Temperature (C)")

```


