---
title: "Atmospheric pCO2"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r parent, child = "/nfs/kryo/work/jenmueller/emlr_cant/utilities/setup_obs.Rmd"}
# this chunk runs the code stored in setup.Rmd
# if required, please refer to instructions given here:
# https://jdblischak.github.io/workflowr/articles/wflow-07-common-code.html
```

```{r define_paths, include=FALSE}
path_atm_pCO2    <- "/nfs/kryo/work/updata/atm_pco2_noaa/"
path_reccap2    <- "/nfs/kryo/work/updata/reccap2/"
path_preprocessing  <- paste(path_root, "/observations/preprocessing/", sep = "")
```

# Read data

- Data source: `Globally averaged marine surface annual mean data` from [ESRL-NOAA](https://www.esrl.noaa.gov/gmd/ccgg/trends/gl_data.html)

```{r read_atm_co2}

co2_atm <- read_table2(
  paste(path_atm_pCO2,
        "co2_annmean_gl.txt",
        sep = ""),
  col_names = FALSE,
  comment = "#"
)

co2_atm <- co2_atm %>% 
  select(-X3)

names(co2_atm) <- c("year", "pCO2")
  
  
```

- Data source:

Global CO2 concentration (ppm)
Prepared by C Le Quéré and M W Jones for the Global Carbon Project, 1 May 2020
This dataset is intended to be used as atmospheric forcing for modelling the evolution of carbon sinks

Data from March 1958 are monthly average from MLO and SPO
provided by NOAA's Earth System Research Laboratory
http://www.esrl.noaa.gov/gmd/ccgg/trends/
When no SPO data are available (including prior to 1975), SPO is constructed from the 1976-2014 average 
MLO-SPO trend and average monthly departure
The last year of data are still preliminary, pending recalibrations of reference gases and other quality control checks.

Data prior to March 1958 are estimated with a cubic spline fit to ice core data from
Joos and Spahni 2008 Rates of change in natural and anthropogenic radiative
forcing over the past 20,000 years PNAS

Annual mean values are calculated for all years.

```{r read_atm_co2_reccap}

co2_atm_reccap2 <- read_table2(
  paste0(path_reccap2,
        "global_co2_merged.txt"),
  col_names = c("year", "pCO2"),
  skip = 16
)

co2_atm_reccap2 <- co2_atm_reccap2 %>% 
  mutate(year = as.integer(round(year))) %>% 
  group_by(year) %>% 
  summarise(pCO2 = mean(pCO2, na.rm = TRUE)) %>% 
  ungroup()

```

# Time series

```{r co2_atm_time_series}

ggplot() +
  # geom_path(data = co2_atm_reccap2, aes(year, pCO2, col="reccap2")) +
  geom_point(data = co2_atm_reccap2, aes(year, pCO2, col="reccap2")) +
  # geom_path(data = co2_atm, aes(year, pCO2, col="NOAA")) +
  geom_point(data = co2_atm, aes(year, pCO2, col="NOAA")) +
  scale_color_brewer(palette = "Set1") +
  theme(legend.title = element_blank())

```

# Write clean file

```{r write_co2_atm_clean_file}

co2_atm %>%
  write_csv(paste(path_preprocessing,
                  "co2_atm.csv",
                  sep = ""))

co2_atm_reccap2 %>%
  write_csv(paste(path_preprocessing,
                  "co2_atm_reccap2.csv",
                  sep = ""))

```
